Accessing the nucleon tensor structure in inclusive deep inelastic scattering

### 2016-07-26 1607 - JMC for g2/Paper/JMC_dis.tex
fixed some references

It is important to notice that $\xi_2=1$ exactly due to CPT invariance \cite{Weinberg-book}, while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of spectral functions we may expect \cite{Accardi:2008ne)} $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2 (\mu^2) \theta (\mu^2-m_\pi^2)$ with the continuum starting at $m_\pi$, the mass of the pion, due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore expect 

@@

Rather than directly using the definition \eqref{e:xi_A}, it is convenient to calculate the inclusive cross section as an integral of the semi-inclusive one, utilize the QCD equation of motions and furthermore summed over all hadron flavors, and take advantage of 
\begin{align}
  \label{eq:SIDIS_to_DIS}
  \sum_h \int \frac{d^3p_h}{(2\pi)2E_h} \Delta^h(l,p_h) = \Xi(l) \ , 
\end{align}
where $\Delta^h$ is the quark fragmentation correlator for production of a hadron of flavor $h$ and momentum $p_h$ \cite{New_testament}. In terms of the TMD fragmentation functions we are interested in, this reads
\begin{align}
  \label{eq:SIDIS_to_DIS_TMDlevel}
  \sum_h \int dz d^2p_{hT} z D_1^h(z,p_{hT}) & = \xi_2 = 1   \\
  \sum_h \int dz d^2p_{hT} E(z,p_{hT}) & = \xi_1 \ ,
\end{align}
where $D_1^h(z,p_{hT})$ is the twist-2 quark fragmentation function as a function of the hadron's collinear momentum fraction $z$ and transverse momentum $p_{hT}$, and $\tilde E^h(z,p_{hT})$ is a chiral-odd twist-3 function defined in \cite{New-testament}.

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\todo{[AA] I am using the notation in Piet Mulder's lecture notes - this will need to be checked.}
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 x_b g_T(x_B) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(x_B) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
Here and we reintroduced the quark flavor $q$ for maximum clarity, and $e_q$ its electric charge.
The first term can be easily integrated with the help of the sum rule \eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a consequence of the QCD equations of motion \cite{New-testament}, then utilize the sum rule \eqref{eq:SIDIS_to_DIS}:
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a non perturbative generalization of the commonly used $\int\tilde E =0$ sum rule introduced in \cite{Jaffe-Ji}. Indeed, calculating the jet correlator
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the new term would vanish.

@@

It is important to estimate the size of teh variou scontributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{tw3} & = \frac{1}{2}\,\sum_a e_a^2
    \biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
&
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_a e_a^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(x),
&
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_a e_a^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x}. 
\end{align} 
These are comapred in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function obtained in the very recent JAM15 fit of polarized DIS asymmetries \cite{JAM15}, that includes a large amount of precise data at large $x$ from Jefferson Lab, and simultaneously fits the higher-twist components of in $g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution, $g_2^{tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix elements, we show a model calculation by Braun et al. \cite{Braun-et-al}; for other estimates, see \cite{Other-tw3-calcs}. To estimate the contributions from quark ($g_2^q$) and jet mass ($g_2^{jet}$) effects, that depend on chiral odd quark-quark matrix elements, we use the recent Pavia15 fit of the transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable also to other
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

@@

\begin{itemize}
\item smaller-$x$: constrains the small-$x$ behavior of transversity.
\end{itemize}
Some ideas to check:
\begin{itemize}
\item JAM13 neutron has almost 0 twist 3, with small positive contribution at small x. Can we use this to highlight teh $g_2^{jet}$ contribution? Or maybe as a cnstrain: if Braun is right then $g_2^{jet}(n) \approx -g_2^{tw3}(n)$. (still, does not help much for the proton).
\item In Jam 13, Fig. 6, it looks like in general the violation of teh BC sum rule is a small-x effect --> constraints on h(x) fits? This cannot really be checked vs. JAM15, that assumes $g_2^(0)$ respects BC. 
\end{itemize}

\begin{figure}[tbh]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left) and neutron (right) $g_2$ structure function compared to the JAM15 fit of the  $g_2-g_2^{\text{WW}}$ (solid black) \cite{JAM15}. The quark and jet contributions are shown with a dotted red and a dot-dashed green line respectively, with uncertainty bands coming form the Pavia15 fit of the transversity function \cite{Pavia15}. The unceratainty in the choice $m_q=5$ GeV and $M_q=100$ GeV is not shown. The pure twist-3 contribution calculated by Braun et al. \cite{Braun-et-al} is shown as a dashed blue line (no uncertainty estiamte was provided in teh original reference).
}
\end{center}
\end{figure}

@@

The zero-th moment corresponds to an extension of teh Burkhardt-Cottingham sum rule \cite{BC,Jaffe-lectures}, that appears to be broken by the non-perturbative spin-flip contribution from the jet function: 
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1(x) \ .
\end{align}
This sum rule, in which the pure twist-3 part does not take part, is a consequence of the Lorentz invariance of the DIS cross section, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. It explicitly displays a contribution from spin-flip processes, that are included in the original derivation by Burkhardt and Cottingham \cite{BC} but do not show up in treatments that only consider free field quark propagators for the struck quark \cite{Jaffe-lectures}.

Assuming the Burkhardt-Cottingham sum rule to be strictly valid, we obtain a strong constraint on the transversity function,
\begin{align}
   \int dx\, \frac{1}{x} h_1(x) = 0 \ .
\end{align}
Even if the BC sun rule is broken by a 
$J = 0$ fixed pole with non-polynomial residue, i.e., if $g_2$ integrates to a finite but non-zero number \cite{Jaffe-lectures}), we obtain that $h_1(x)/x$ must be integrable. This entails a bound on the small $x$ behavior of the transversity,
\begin{align}
  h_1(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\end{align}
This bound will be very useful, \eg, in transversity fits, where the data at small $x$ is as yet very limited.

>>>>

It is important to notice that $\xi_2=1$ exactly due to CPT invariance \cite{Weinberg-book}, while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of spectral functions we may expect \cite{Accardi:2008ne} $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2 (\mu^2) \theta (\mu^2-m_\pi^2)$ with the continuum starting at $m_\pi$, the mass of the pion, due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore expect 

@@

Rather than directly using the definition \eqref{e:xi_A}, it is convenient to calculate the inclusive cross section as an integral of the semi-inclusive one, utilize the QCD equation of motions and furthermore summed over all hadron flavors, and take advantage of 
\begin{align}
  \label{eq:SIDIS_to_DIS}
  \sum_h \int \frac{d^3p_h}{(2\pi)2E_h} \Delta^h(l,p_h) = \Xi(l) \ , 
\end{align}
where $\Delta^h$ is the quark fragmentation correlator for production of a hadron of flavor $h$ and momentum $p_h$ \cite{Bacchetta:2006tn}. In terms of the TMD fragmentation functions we are interested in, this reads
\begin{align}
  \label{eq:SIDIS_to_DIS_TMDlevel}
  \sum_h \int dz d^2p_{hT} z D_1^h(z,p_{hT}) & = \xi_2 = 1   \\
  \sum_h \int dz d^2p_{hT} E(z,p_{hT}) & = \xi_1 \ ,
\end{align}
where $D_1^h(z,p_{hT})$ is the twist-2 quark fragmentation function as a function of the hadron's collinear momentum fraction $z$ and transverse momentum $p_{hT}$, and $\tilde E^h(z,p_{hT})$ is a chiral-odd twist-3 function defined in \cite{Bacchetta:2006tn}.

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\todo{[AA] I am using the notation in Piet Mulder's lecture notes - this will need to be checked.}
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 x_b g_T(x_B) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(x_B) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
Here and we reintroduced the quark flavor $q$ for maximum clarity, and $e_q$ its electric charge.
The first term can be easily integrated with the help of the sum rule \eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then utilize the sum rule \eqref{eq:SIDIS_to_DIS}:
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the new term would vanish.

@@

It is important to estimate the size of teh variou scontributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{tw3} & = \frac{1}{2}\,\sum_a e_a^2
    \biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
&
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_a e_a^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(x),
&
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_a e_a^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x}. 
\end{align} 
These are comapred in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function
obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $x$
from Jefferson Lab, and simultaneously fits the higher-twist components of in
$g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution,
$g_2^{tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a model calculation by Braun et al. \cite{Braun-et-al}; for
other estimates, see \cite{Other-tw3-calcs}. To estimate the contributions
from quark ($g_2^q$) and jet mass ($g_2^{jet}$) effects, that depend on chiral
odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

@@

\begin{itemize}
\item smaller-$x$: constrains the small-$x$ behavior of transversity.
\end{itemize}
Some ideas to check:
\begin{itemize}
\item JAM13 neutron has almost 0 twist 3, with small positive contribution at small x. Can we use this to highlight teh $g_2^{jet}$ contribution? Or maybe as a cnstrain: if Braun is right then $g_2^{jet}(n) \approx -g_2^{tw3}(n)$. (still, does not help much for the proton).
\item In Jam 13, Fig. 6, it looks like in general the violation of the BC sum rule is a small-x effect --> constraints on h(x) fits? This cannot really be checked vs. JAM15, that assumes $g_2^(0)$ respects BC. 
\end{itemize}

\begin{figure}[tbh]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure function compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The unceratainty in the choice
$m_q=5$ GeV and $M_q=100$ GeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estiamte was provided in teh original reference). 
}
\end{center}
\end{figure}

@@

The zero-th moment corresponds to an extension of teh Burkhardt-Cottingham sum rule \cite{Burkhardt:1970ti,Jaffe:1996zw}, that appears to be broken by the non-perturbative spin-flip contribution from the jet function: 
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1(x) \ .
\end{align}
This sum rule, in which the pure twist-3 part does not take part, is a consequence of the Lorentz invariance of the DIS cross section, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. It explicitly displays a contribution from spin-flip processes, that are included in the original derivation by Burkhardt and Cottingham \cite{Burkhardt:1970ti} but do not show up in treatments that only consider free field quark propagators for the struck quark \cite{Jaffe:1996zw}.

Assuming the Burkhardt-Cottingham sum rule to be strictly valid, we obtain a strong constraint on the transversity function,
\begin{align}
   \int dx\, \frac{1}{x} h_1(x) = 0 \ .
\end{align}
Even if the BC sun rule is broken by a 
$J = 0$ fixed pole with non-polynomial residue, i.e., if $g_2$ integrates to a finite but non-zero number \cite{Jaffe:1996zw}), we obtain that $h_1(x)/x$ must be integrable. This entails a bound on the small $x$ behavior of the transversity,
\begin{align}
  h_1(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\end{align}
This bound will be very useful, \eg, in transversity fits, where the data at small $x$ is as yet very limited.
